---
title: 'Class 11: Scraping Multiple Webpages'
author: "Dominic Sanchez"
date: "2026-09-30"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_download: true
    code_folding: show
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Learning Objectives

- Enable you to scrape multiple webpages (from the same site), using a `for` loop type approach

- Enable you to scrape multiple webpages (from the same site), using the [purrr package](https://purrr.tidyverse.org/),
which is a functional programming approach
*Note:* In this process, we will create our own function


## Re-examining the bulletin example

Let us try to expand on the example from last class to include the following:
[ACC](https://bulletin.miamioh.edu/courses-instruction/acc/),
[BUS](https://bulletin.miamioh.edu/courses-instruction/bus/),
[ISA](https://bulletin.miamioh.edu/courses-instruction/isa/),
[MKT](https://bulletin.miamioh.edu/courses-instruction/mkt/)


**COMMENT:** Obviously, you can repeat your code few times with changing the URL. This will 
work for a few links, but would be quite tedious for large amounts of scraping. We will follow the 
**DRY Principle**, which stands for **Don't Repeat Yourself**.


```{r bulletin}
dept_abbrv = c('acc', 'bus', 'isa', 'mkt')

courses_df = data.frame()

for (dept in dept_abbrv){
  cat(rep("-", 50), "\n")
  print(dept)
  
  
  bulletin_url = paste0(
    'https://bulletin.miamioh.edu/courses-instruction/', dept, '/'
  )
  
  webpage = rvest::read_html(bulletin_url)
  
  webpage |> rvest::html_elements("strong") |> rvest::html_text2() -> titles
  
  webpage |> rvest::html_elements("p.courseblockdesc") |> rvest::html_text2() -> descs
  #webpage |> rvest::html_elemts("#sc_sccoursedescs > div > p.courseblockdesc") |> rvest::html_text2() -> descs
  
  df = data.frame(title = titles, desc = descs)
  courses_df = rbind(courses_df, df) # similar to append in which it stacks the new dataframe on top of the accumulated one
}
```



## `purrr` approach for the same problem

This is a functional programming approach, and we will write our own function. In R, we use the
function called `function` to create a function (def in python). As opposed to Python, R functions return a single objects.

```{r purrr}
dept_abbrv = c('acc', 'bus', 'isa', 'mkt')

# This function takes an abbreviastion as an input and returns a data_frame as an output
scrape_mu_bulletin = function(dept){
  bulletin_url = paste0(
    'https://bulletin.miamioh.edu/courses-instruction/', dept, '/'
  )
  
  webpage = rvest::read_html(bulletin_url)
  
  webpage |> rvest::html_elements("p.courseblocktitle") |> rvest::html_text2() -> titles
  
  webpage |> rvest::html_elements("p.courseblockdesc") |> rvest::html_text2() -> descs
  
  df = data.frame(title = titles, desc = descs)
  return(df)
}

test = scrape_mu_bulletin(dept = 'eco')

# to use functions from purrr, we will use a map_*(), or map2_*(), or a pmap_*(); * = something
# map uses a function that takes a single input (or you are only changing a single input)
# map2 uses a function that takes two inputs
# pmap >= 3
# after the _, you put the type of object you are getting back

# the .x is the vector that we will iterate over
# and the .f is the function that we will use


courses_df_from_purrr = purrr::map_df(.x = dept_abbrv, .f = scrape_mu_bulletin)
```


## Scrape data from the website planecrashinfo

The website that we want to scrape is called: <https://www.planecrashinfo.com/> and it contains a
database of plane crashes by year.


```{r plane_crashes}
year = 2025

crash_url = paste0('https://www.planecrashinfo.com/', year, '/', year, '.htm')

crash_page = rvest::read_html(crash_url)
```




